---
title: 'Metagene around MS Gencode PROMPT 3p end'
author: "Manfred Schmid"
output: 
  pdf_document:
    toc: true 
    toc_depth: 3
    fig_caption: true
---

`r format(Sys.time(), "%d %B, %Y; %R")`

## Setup

```{r setup, echo=TRUE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, 
                      fig.path=paste0('../../Figures/MS_Gencode_PROMPT_3endseq_metagene_withsizefactors/'), 
                      dev='pdf', 
                      echo=TRUE, warning=FALSE, message=FALSE, 
                      error=TRUE)
```

```{r load packages, echo=T, warning=F, message=F}
suppressWarnings(library('tidyverse'))
suppressWarnings(library('magrittr'))
suppressWarnings(library('knitr'))
suppressWarnings(library('broom'))
suppressWarnings(library('RMetaTools'))
```


```{r}
lineplot_style <- list(
  scale_color_manual(values=c('#1f77b4', '#ff7f0e', '#2ca02c', 'red', 'navyblue'), na.value='lightgray'),
  scale_fill_manual(values=c('#1f77b4', '#ff7f0e', '#2ca02c', 'red', 'navyblue'), na.value='lightgray'),
  theme_bw(),
  theme(axis.text.x = element_text(angle=45, hjust=1),
        panel.grid = element_blank(),
        panel.background = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line()))


areaplot_style <- list(
  scale_fill_manual(values=c('#2C3E95', '#4EAE48', '#397DB9'), na.value='lightgray'),
  scale_color_manual(values=c('#2C3E95', '#4EAE48', '#397DB9'), na.value='lightgray'),
  theme_bw(),
  theme(axis.text.x = element_text(angle=45, hjust=1),
        panel.grid = element_blank(),
        panel.background = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line()))
```

#### deeptools run

```{bash, eval=FALSE}
#!/bin/sh
##cd /home/schmidm/faststorage/Guifen/Lexogen_July_Sept_Nov/scripts/deeptools/MS_Gencode_PROMPTs/
##sbatch --account=thj_common --mem=2g deeptools_allGuifen_TESm1p1kb_offcenter.sh

. /home/schmidm/miniconda2/etc/profile.d/conda.sh
conda activate deeptools3
cd /home/schmidm/faststorage/Guifen/Lexogen_July_Sept_Nov/scripts/deeptools/MS_Gencode_PROMPTs

anno="/home/schmidm/annotations/hg38/Gencode_v28/PROMPT_TSS_to5kb_TSS_toMaxIn5kb_withend.bed"
#awk '{OFS="\t"}{if($6 == "+"){$2=($3-1025);$3+=1025;print $0}}' $anno > ${anno/.bed/_TESpm1025_plus.bed}
#awk '{OFS="\t"}{if($6 == "-"){$3=($2+1025);$2-=1025;print $0}}' $anno > ${anno/.bed/_TESpm1025_minus.bed}


outdir="/home/schmidm/faststorage/Guifen/Lexogen_July_Sept_Nov/scripts/deeptools/MS_Gencode_PROMPTs/deeptools_out/all_TESm1025p1025_3pseq_all_deeptools_out"
mkdir $outdir
sizeFactors="/home/schmidm/faststorage/Guifen/Lexogen_July_Sept_Nov/last_exon_sizeFactors.table"

cat $sizeFactors | sed 1d | while read line;
#tail -1 $sizeFactors | while read line;
#grep siZCCHC8 $sizeFactors | grep ZFC3H1 | while read line;
do
  arr=($line)
  fname=${arr[0]/./+}
  month=${arr[1]}
  echo "${month} ${fname}"

  bw_plus="/home/schmidm/faststorage/Guifen/${arr[5]}/bigwig/${fname}_plus.bw"
  bw_minus=${bw_plus//plus.bw/minus.bw}

  name="raw_${month}_${fname}"
  mat="${outdir}/${name}.gz"

  echo $mat

  qx python /home/schmidm/ms_tools/MS_Metagene_Tools/computeMatrixStranded.pyc scale-regions -Rp ${anno/.bed/_TESpm1025_plus.bed} -Rm ${anno/.bed/_TESpm1025_minus.bed} -Sp $bw_plus -Sm $bw_minus -o $mat -a 0 -b 0 --binSize=50 --regionBodyLength 2050 --missingDataAsZero --averageTypeBins sum --quiet
done

echo "ALL DONE"

```


## load and R 

#### load PROMPT classification

```{r}
(prompt_info <- read_tsv('../../data/PROMPT_sensitivities_Z1_Z8_RRP40_withextraclasses_withsizeFactors.txt'))
```


#### load metagene matrices

```{r, eval=FALSE}
path <- '/Volumes/GenomeDK/faststorage/Guifen/Lexogen_July_Sept_Nov/scripts/deeptools/MS_Gencode_PROMPTs/deeptools_out/all_TESm1025p1025_3pseq_all_deeptools_out/'

(fnames <- dir(path) %>% keep(grepl('^raw.*.gz', .) & !grepl('RBM7', .)))
```


```{r, eval=FALSE}
(df <- lapply(fnames, function(fname) RMetaTools::load_deeptoolsmatrix3(paste0(path,fname)) %>%
         mutate(sample_name = sub('.*raw_', '', fname) %>% sub('.gz', '', .))) %>%
         bind_rows)
```

```{r, eval=FALSE}
(df %<>%
  mutate(id = sub('center_of:', '', id),
         rel_pos = rel_pos - 1000) %>%
  left_join(., prompt_info) %>%
  separate(sample_name, c('series', 'siRNA', 'PAP', 'fraction', 'rep'), sep='_')  %>%
  mutate(series = ifelse(is.na(rep), series, paste0('Novx', rep))) %>%
  dplyr::select(class3, class5, id, rel_pos, series, siRNA, PAP, fraction, value))
```


#### load size factors

```{r, eval=FALSE}
(nfs <- read_tsv('../../data/last_exon_sizeFactors.table'))
```

```{r, eval=FALSE}
(nfs %<>%
  tidyr::separate(lib_name, c('siRNA', 'PAP', 'fraction', 'rep'), sep='_') %>%
   mutate(series = ifelse(is.na(rep), series, paste0('Novx',rep)),
          siRNA = sub('\\.', '\\+', siRNA)) %>%
   dplyr::select(-bg_bam_dir, -lib, -rep))
```

#### normalize using size factors

```{r, eval=FALSE}
(df %<>% left_join(., nfs) %>%
  mutate(value = value/geom_mean_sf,
         class3 = factor(class3, levels=c('PAXT', 'NEXT', 'BOTH')),
         class5 = factor(class5, levels=c('PAXT', 'NEXT', 'BOTH', 'NEXTonly', 'PAXTonly', 'not sig')),
         siRNA = factor(siRNA, levels=c('siGFP', 'siRRP40', 'siZCCHC8', 'siZFC3H1', 'siZCCHC8+ZFC3H1'))))
```

```{r, eval=FALSE}
save(df, file = '../../data/MS_Gencode_PROMPTs_3pseq_metagene.RData')
```

```{r, eval=TRUE}
load('../../data/MS_Gencode_PROMPTs_3pseq_metagene.RData')
df
```

```{r}
(prompt_info <- read_tsv('../../data/PROMPT_sensitivities_Z1_Z8_RRP40_withextraclasses_withsizeFactors.txt'))
```


```{r}
df %<>%
  dplyr::select(-class3, -class5) %>%
  left_join(., distinct(prompt_info, id, class3, class5)) %>%
  mutate(class3 = factor(class3, levels=c('PAXT', 'NEXT', 'BOTH')),
         class5 = factor(class5, levels=c('PAXT', 'NEXT', 'BOTH', 'NEXTonly', 'PAXTonly', 'not sig')))
```


#### prepare for metagene


```{r}
df %<>%
  mutate(log2_value = log2(value+1))

ctrls <- df %>%
  filter(siRNA == "siGFP") %>%
  dplyr::rename(ctrl_log2_value = log2_value) %>%
  dplyr::select(id, rel_pos, series, PAP, fraction, ctrl_log2_value)

log2FCs <- df %>%
  filter(siRNA != "siGFP") %>%
  left_join(., ctrls) %>%
  mutate(log2FC = log2_value - ctrl_log2_value,
         siRNA = factor(siRNA, levels=c('siRRP40', 'siZCCHC8', 'siZFC3H1', 'siZCCHC8+ZFC3H1')))
```

## violin plots


```{r}
per_id <- df %>%
  group_by(id, class3, series, siRNA, PAP, fraction) %>%
  summarize(total_signal = sum(value)) %>%
  group_by(id, class3, siRNA, PAP, fraction) %>%
  summarize(total_signal = mean(total_signal))

ctrls <- per_id %>%
  filter(siRNA == "siGFP") %>%
  dplyr::rename(ctrl_total_signal = total_signal) %>%
  ungroup %>%
  dplyr::select(id, PAP, fraction, ctrl_total_signal)

log2FC_per_id <- per_id %>%
  ungroup %>%
  filter(siRNA != "siGFP") %>%
  left_join(., ctrls) %>%
  mutate(log2FC = log2(total_signal+1) - log2(ctrl_total_signal+1),
         siRNA = factor(siRNA, levels=c('siRRP40', 'siZCCHC8', 'siZFC3H1', 'siZCCHC8+ZFC3H1')))
```


```{r violins log2FCs TES pm 1kb}
ggplot(log2FC_per_id, aes(x=siRNA, y=log2FC, fill=siRNA)) +
  geom_violin() +
  geom_boxplot(width=.2, fill='lightgray', outlier.shape=NA) +
  facet_grid(fraction+PAP~class3) +
  geom_hline(yintercept = 0, linetype=2) +
  theme_bw() +
  theme(panel.grid=element_blank())
```



## metagene raw values profiles

#### 2-class version

```{r}
(meta_values3 <- df %>%
  filter(!is.na(class3)) %>%
  group_by(class3, siRNA, PAP, fraction, rel_pos) %>%
  do(tidy(t.test(.$log2_value))))
```


```{r cluster metagene profile normalized values 2classes}
meta_values3 %>%
  filter(class3 != 'BOTH', class3 != 'not sig') %>%
  ggplot(. ,aes(x=rel_pos, y=estimate)) +
  geom_line(aes(color=class3)) +
  geom_ribbon(aes(ymin=conf.low, ymax=conf.high, fill=class3), alpha=.2)  +
  facet_grid(fraction+PAP~siRNA)  +
  ylab('Average Log2 Signal') +
  lineplot_style
```


```{r}
(meta_log2FCs3 <- log2FCs %>%
  group_by(class3, siRNA, PAP, fraction, rel_pos) %>%
  do(tidy(t.test(.$log2FC))))
```


```{r cluster metagene profile log2FC 2classes}
meta_log2FCs3 %>%
  filter(class3 != 'BOTH', class3 != 'not sig') %>%
  ggplot(. ,aes(x=rel_pos, y=estimate)) +
  geom_line(aes(color=class3)) +
  geom_ribbon(aes(ymin=conf.low, ymax=conf.high, fill=class3), alpha=.2)  +
  facet_grid(fraction+PAP~siRNA)  +
  ylab('Average Log2 Signal') +
  lineplot_style
```


#### 3-class version


```{r cluster metagene profile normalized values 3classes}
meta_values3 %>%
  filter(class3 != 'not sig') %>%
  ggplot(. ,aes(x=rel_pos, y=estimate)) +
  geom_line(aes(color=class3)) +
  geom_ribbon(aes(ymin=conf.low, ymax=conf.high, fill=class3), alpha=.2)  +
  facet_grid(fraction+PAP~siRNA)  +
  ylab('Average Log2 Signal') +
  lineplot_style
```

```{r PROMPT metagene profile log2FC 3classes}
meta_log2FCs3 %>%
  filter(class3 != 'not sig') %>%
  ggplot(. ,aes(x=rel_pos, y=estimate)) +
  geom_line(aes(color=class3)) +
  geom_ribbon(aes(ymin=conf.low, ymax=conf.high, fill=class3), alpha=.2)  +
  facet_grid(fraction+PAP~siRNA)  +
  ylab('Average Log2 Signal') +
  lineplot_style
```

```{r PROMPT metagene profile log2FC 3classes areas}
meta_log2FCs3 %>%
  filter(class3 != 'not sig', siRNA != 'siRRP40') %>%
  ungroup %>%
  mutate(siRNA = factor(siRNA, levels=c('siZCCHC8+ZFC3H1', 'siZCCHC8', 'siZFC3H1'))) %>%
  ggplot(.) +
  geom_ribbon(aes(x=rel_pos, ymin=0, ymax=estimate, fill=siRNA), alpha=.75) +
  geom_line(aes(x=rel_pos, y=estimate, color=siRNA)) +
  #geom_ribbon(aes(ymin=conf.low, ymax=conf.high, fill=class3), alpha=.2)  +
  facet_grid(fraction+PAP~class3)  +
  ylab('Average Log2 Signal') +
  areaplot_style
```





#### 4-class version

```{r}
(meta_values4 <- df %>%
  mutate(log2_value = log2(value+1),
         class4 = ifelse(as.character(class5) != 'PAXTonly', as.character(class5), 'PAXT')) %>%
  group_by(class4, siRNA, PAP, fraction, rel_pos) %>%
  do(tidy(t.test(.$log2_value))) %>%
   ungroup %>%
   mutate(class4 = factor(class4, levels=c('PAXT', 'BOTH', 'NEXT', 'NEXTonly', 'not sig'))))
```


```{r cluster metagene profile normalized values 4classes}
meta_values4 %>%
  filter(class4 != 'not sig') %>%
  ggplot(. ,aes(x=rel_pos, y=estimate)) +
  geom_line(aes(color=class4)) +
  geom_ribbon(aes(ymin=conf.low, ymax=conf.high, fill=class4), alpha=.2)  +
  facet_grid(fraction+PAP~siRNA)  +
  ylab('Average Log2 Signal') +
  lineplot_style
```


```{r}
(meta_log2FCs4 <- log2FCs %>%
  mutate(class4 = ifelse(as.character(class5) != 'PAXTonly', as.character(class5), 'PAXT')) %>%
  group_by(class4, siRNA, PAP, fraction, rel_pos) %>%
  do(tidy(t.test(.$log2FC))) %>%
   ungroup %>%
   mutate(class4 = factor(class4, levels=c('PAXT', 'BOTH', 'NEXT', 'NEXTonly', 'not sig'))))
```


```{r cluster metagene profile log2FC 4classes}
meta_log2FCs4 %>%
  filter(class4 != 'not sig') %>%
  ggplot(. ,aes(x=rel_pos, y=estimate)) +
  geom_line(aes(color=class4)) +
  geom_ribbon(aes(ymin=conf.low, ymax=conf.high, fill=class4), alpha=.2)  +
  facet_grid(fraction+PAP~siRNA)  +
  ylab('Average Log2 Signal') +
  lineplot_style
```


```{r PROMPT metagene profile log2FC 4classes areas}
meta_log2FCs4 %>%
  filter(class4 != 'not sig', siRNA != 'siRRP40') %>%
  ungroup %>%
  mutate(siRNA = factor(siRNA, levels=c('siZCCHC8+ZFC3H1', 'siZCCHC8', 'siZFC3H1'))) %>%
  ggplot(.) +
  geom_ribbon(aes(x=rel_pos, ymin=0, ymax=estimate, fill=siRNA), alpha=.75) +
  geom_line(aes(x=rel_pos, y=estimate, color=siRNA)) +
  #geom_ribbon(aes(ymin=conf.low, ymax=conf.high, fill=class3), alpha=.2)  +
  facet_grid(fraction+PAP~class4)  +
  ylab('Average Log2 Signal') +
  areaplot_style
```




#### 5-class version

```{r}
(meta_values5 <- df %>%
  mutate(log2_value = log2(value+1)) %>%
  group_by(class5, siRNA, PAP, fraction, rel_pos) %>%
  do(tidy(t.test(.$log2_value))))
```


```{r cluster metagene profile normalized values 5classes}
meta_values5 %>%
  filter(class5 != 'not sig') %>%
  ggplot(. ,aes(x=rel_pos, y=estimate)) +
  geom_line(aes(color=class5)) +
  geom_ribbon(aes(ymin=conf.low, ymax=conf.high, fill=class5), alpha=.2)  +
  facet_grid(fraction+PAP~siRNA)  +
  ylab('Average Log2 Signal') +
  lineplot_style
```


```{r}
(meta_log2FCs5 <- log2FCs %>%
  group_by(class5, siRNA, PAP, fraction, rel_pos) %>%
  do(tidy(t.test(.$log2FC))))
```


```{r cluster metagene profile log2FC 5classes}
meta_log2FCs5 %>%
  filter(class5 != 'not sig') %>%
  ggplot(. ,aes(x=rel_pos, y=estimate)) +
  geom_line(aes(color=class5)) +
  geom_ribbon(aes(ymin=conf.low, ymax=conf.high, fill=class5), alpha=.2)  +
  facet_grid(fraction+PAP~siRNA)  +
  ylab('Average Log2 Signal') +
  lineplot_style
```


```{r PROMPT metagene profile log2FC 5classes areas}
meta_log2FCs5 %>%
  filter(class5 != 'not sig', siRNA != 'siRRP40') %>%
  ungroup %>%
  mutate(siRNA = factor(siRNA, levels=c('siZCCHC8+ZFC3H1', 'siZCCHC8', 'siZFC3H1'))) %>%
  ggplot(.) +
  geom_ribbon(aes(x=rel_pos, ymin=0, ymax=estimate, fill=siRNA), alpha=.75) +
  geom_line(aes(x=rel_pos, y=estimate, color=siRNA)) +
  #geom_ribbon(aes(ymin=conf.low, ymax=conf.high, fill=class3), alpha=.2)  +
  facet_grid(fraction+PAP~class5)  +
  ylab('Average Log2 Signal') +
  areaplot_style
```



## with vs without hexamer at TES


```{r}
(hexamer_info <- rtracklayer::import('../../data/MS_Gencode_PROMPTs_with_class3_class5_hexameratTES.bed') %>%
  data.frame %>%
  tbl_df %>%
  tidyr::separate(name, c('id', 'class3', 'class5', 'hexamer'), sep=':') %>%
  dplyr::select(id, hexamer))
```


```{r }
(meta_log2FCs3 <- log2FCs %>%
   left_join(., hexamer_info) %>%
  group_by(class3, hexamer, siRNA, PAP, fraction, rel_pos) %>%
  do(tidy(t.test(.$log2FC))))


```

```{r cluster metagene profile log2FC 2classes split by hexamer}
meta_log2FCs3 %>%
  filter(class3 != 'not sig', class3 != 'BOTH') %>%
  ggplot(. ,aes(x=rel_pos, y=estimate)) +
  geom_line(aes(color=class3, linetype=hexamer)) +
  #geom_ribbon(aes(ymin=conf.low, ymax=conf.high, fill=class3, alpha=hexamer), alpha=.2)  +
  facet_grid(fraction+PAP~class3+siRNA)  +
  ylab('Average Log2 Signal') +
  lineplot_style
```


```{r cluster metagene profile log2FC 5classes split by hexamer}
(meta_log2FCs5 <- log2FCs %>%
   left_join(., hexamer_info) %>%
  group_by(class5, hexamer, siRNA, PAP, fraction, rel_pos) %>%
  do(tidy(t.test(.$log2FC))))

meta_log2FCs5 %>%
  filter(class5 != 'not sig') %>%
  ggplot(. ,aes(x=rel_pos, y=estimate)) +
  geom_line(aes(color=class5)) +
  geom_ribbon(aes(ymin=conf.low, ymax=conf.high, fill=class5), alpha=.2)  +
  facet_grid(fraction+PAP~siRNA+hexamer)  +
  ylab('Average Log2 Signal') +
  lineplot_style
```



#### significance of +/- hexamer

which position to interrogate
```{r cluster metagene profile log2FC 2classes split by hexamer center zoom}
meta_log2FCs3 %>%
  filter(class3 == 'PAXT' | class3 == 'NEXT',
         rel_pos >= -100 & rel_pos <= 100) %>%
  ggplot(. ,aes(x=rel_pos, y=estimate)) +
  geom_line(aes(color=class3, linetype=hexamer)) +
  #geom_ribbon(aes(ymin=conf.low, ymax=conf.high, fill=class3, alpha=hexamer), alpha=.2)  +
  facet_grid(fraction+PAP~class3+siRNA)  +
  ylab('Average Log2 Signal') +
  lineplot_style
```

--> by far most signals centers at exact 0

--> do a MWU test for position 0 H0 being that hexamer and no-hexamer classes are equal within each type

```{r}
(log2FCs_with_hex <- log2FCs %>%
  filter(class3 == 'PAXT' | class3 == 'NEXT',
         rel_pos == 0) %>%
   left_join(., hexamer_info))
```

#### two-sided test

```{r}
log2FCs_with_hex %>%
  group_by(class3, siRNA, PAP, fraction) %>%
  do(tidy(wilcox.test(.$log2FC~.$hexamer))) %>%
  dplyr::select(class3, siRNA, PAP, fraction, p.value) %>%
  spread(class3, p.value) %>%
  kable
```

#### one-sided tests


```{r}
levels(factor(log2FCs_with_hex$hexamer))
```

```{r}
log2FCs_with_hex %>%
  group_by(class3, siRNA, PAP, fraction) %>%
  do(tidy(wilcox.test(.$log2FC~.$hexamer, alternative='greater'))) %>%
  dplyr::select(class3, siRNA, PAP, fraction, p.value) %>%
  spread(class3, p.value) %>%
  kable(., digits = 20)
```

```{r}
log2FCs_with_hex %>%
  group_by(class3, siRNA, PAP, fraction) %>%
  do(tidy(wilcox.test(.$log2FC~.$hexamer, alternative='less'))) %>%
  dplyr::select(class3, siRNA, PAP, fraction, p.value) %>%
  spread(class3, p.value) %>%
  kable(., digits = 100)
```



## siZ1Z8 vs siZ8 significance

```{r}
z8 <- df %>%
  filter(siRNA == "siZCCHC8") %>%
  dplyr::rename(z8_log2_signal = log2_value) %>%
  dplyr::select(id, rel_pos, series, PAP, fraction, z8_log2_signal)

z1z8 <- df %>%
  filter(siRNA == "siZCCHC8+ZFC3H1") %>%
  dplyr::rename(z1z8_log2_signal = log2_value) %>%
  left_join(., z8) %>%
  mutate(z1z8vsz8_log2FC = z1z8_log2_signal - z8_log2_signal)
```

```{r}
(ttest_meta_z1z8 <- z1z8 %>%
  group_by(class3, PAP, fraction, rel_pos) %>%
  do(tidy(t.test(.$z1z8vsz8_log2FC))))
```

```{r Z1Z8 vs Z8 log2FC}
ttest_meta_z1z8 %>%
ggplot(., aes(x=rel_pos, y=estimate, color=class3)) +
  geom_line() +
  geom_hline(yintercept = 0, linetype=2) +
  facet_grid(fraction+PAP~class3) +
  theme_bw() +
  theme(panel.grid=element_blank(),
        axis.text.x = element_text(angle=45, hjust=1))
```


```{r}
(wilcox_meta_z1z8 <- z1z8 %>%
  group_by(class3, PAP, fraction, rel_pos) %>%
  do(tidy(wilcox.test(.$z1z8vsz8_log2FC))))
```

```{r}
wilcox_meta_z1z8 %<>%
  group_by(PAP, fraction, class3) %>%
  mutate(FDR = p.adjust(p.value, method='BH')) 
```

```{r Z1Z8 vs Z8 wilcox fdr per line}
wilcox_meta_z1z8 %>%
  filter(!is.na(class3)) %>%
  ungroup %>%
  mutate(class3 = factor(class3, levels = c('BOTH', 'NEXT', 'PAXT'))) %>%
ggplot(., aes(x=rel_pos, y=-log10(FDR), color=class3)) +
  geom_line() +
  coord_cartesian(ylim=c(0,100)) +
  facet_grid(fraction+PAP~class3) +
  geom_hline(yintercept = 2, color='black', linetype = 2) + 
  geom_hline(yintercept = 5, color='black', linetype = 4) + 
  lineplot_style
```


```{r}
wilcox_meta_z1z8 %<>%
  ungroup %>%
  mutate(FDR_global = p.adjust(p.value, method='BH')) 
```


```{r Z1Z8 vs Z8 wilcox fdr global}
wilcox_meta_z1z8 %>%
ggplot(., aes(x=rel_pos, y=-log10(FDR_global), color=class3)) +
  geom_line() +
  coord_cartesian(ylim=c(0,100)) +
  geom_hline(yintercept = c(1,2,3,4,5), color='orange') + 
  facet_grid(fraction+PAP~class3) +
  theme_bw() +
  theme(panel.grid=element_blank(),
        axis.text.x = element_text(angle=45, hjust=1))
```


## sessionInfo
```{r}
sessionInfo()
```

