---
title: 'Analysis and classification of clusters'
author: "Manfred Schmid"
output: 
  pdf_document:
    toc: true 
    toc_depth: 3
    fig_caption: true
editor_options: 
  chunk_output_type: console
---

`r format(Sys.time(), "%d %B, %Y; %R")`

## Setup

```{r setup, echo=TRUE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, 
                      fig.path=paste0('../../Figures/cluster_log2FC/'), 
                      dev='pdf', 
                      echo=TRUE, warning=FALSE, message=FALSE, 
                      error=TRUE)
```

```{r load packages, echo=T, warning=F, message=F}
suppressWarnings(library('tidyverse'))
suppressWarnings(library('magrittr'))
suppressWarnings(library('knitr'))
suppressWarnings(library('broom'))
```


## load the data

```{r}
load('../../data/res_ds_combined_RRP40Z1Z8clusters_lastexonsf.RData', verbose = TRUE)
res_df
```


## classify by NEXT vs PAXT sensitivity

total sig up vs down regulated count:
```{r}
res_df %>%
  distinct(row, sig, comparison, libtype) %>%
  group_by(comparison, sig, libtype) %>%
  summarize(cnt = n()) %>%
  spread(sig, cnt) %>%
  kable
```

--> ZFC3H1 has smallest bias towards upregulation


#### classify using Z1 vs Z8 upregulated

```{r}
z8_up <- res_df %>%
  filter(comparison == 'siZCCHC8_rel_siGFP', sig == 'sig_up') %$%
  row %>%
  unique

z1_up <- res_df %>%
  filter(comparison == 'siZFC3H1_rel_siGFP', sig == 'sig_up') %$%
  row %>%
  unique

both_clusters <- intersect(z8_up, z1_up) %>% data.frame(cluster=.) %>% mutate(class='BOTH')
next_clusters <- setdiff(z8_up, z1_up) %>% data.frame(cluster=.) %>% mutate(class='NEXT')
paxt_clusters <- setdiff(z1_up, z8_up) %>% data.frame(cluster=.) %>% mutate(class='PAXT')
notsig_clusters <- setdiff(unique(res_df$row), c(z8_up, z1_up)) %>% data.frame(cluster=.) %>% mutate(class='NOT_SIG')
```

```{r}
(cluster_classes <- bind_rows(both_clusters, next_clusters, paxt_clusters, notsig_clusters) %>%
  tbl_df)
```

```{r}
table(cluster_classes$class)
```


```{r}
save(cluster_classes, file = '../../data/cluster_classes.RData')
```

```{r}
load('../../data/cluster_classes.RData')
cluster_classes
```

```{r}
res_df %<>%
  left_join(., dplyr::rename(cluster_classes, row=cluster)) %>%
   mutate(class = factor(class, levels=c('PAXT', 'NEXT', 'BOTH', 'NOT_SIG')))
```

#### PAXT examples
```{r}
res_df %>%
  filter(class == 'PAXT', comparison == 'siZFC3H1_rel_siGFP') %>%
  group_by(row) %>%
  arrange(-log2FoldChange)
```


#### NEXT examples
```{r}
res_df %>%
  filter(class == 'NEXT', comparison == 'siZCCHC8_rel_siGFP') %>%
  group_by(row) %>%
  arrange(padj)
```


#### BOTH examples
```{r}
res_df %>%
  filter(class == 'BOTH', comparison == 'siZCCHC8.ZFC3H1_rel_siGFP') %>%
  group_by(row) %>%
  arrange(padj)
```



## scatter Z1 vs Z8 with class per lib type

```{r}
lfc_spread <- res_df %>%
  dplyr::select(row, comparison, libtype, class, log2FoldChange) %>%
  spread(comparison, log2FoldChange)
```

```{r}
ns_lfc_spread <- filter(lfc_spread, class == 'NOT_SIG')
sig_lfc_spread <- filter(lfc_spread, class != 'NOT_SIG')
```


```{r LFCs Z1 vs Z8 with classes colored original}
ggplot(sample_n(ns_lfc_spread,20000),
       aes(x=siZCCHC8_rel_siGFP, y=siZFC3H1_rel_siGFP)) +
  geom_point(color = 'black', alpha=.01, size=.01) +
  geom_point(data=sig_lfc_spread, aes(x=siZCCHC8_rel_siGFP, y=siZFC3H1_rel_siGFP, color=class), size=.1, alpha=.1) +
  facet_wrap(~libtype, scales='free') +
  theme_bw() +
  theme(panel.grid = element_blank())
```


```{r LFCs Z1 vs Z8 with classes colored original floored}
sig_lfc_spread %>%
  mutate(siZCCHC8_rel_siGFP = ifelse(siZCCHC8_rel_siGFP > 10, 10, siZCCHC8_rel_siGFP),
         siZCCHC8_rel_siGFP = ifelse(siZCCHC8_rel_siGFP < -10, -10, siZCCHC8_rel_siGFP),
         siZFC3H1_rel_siGFP = ifelse(siZFC3H1_rel_siGFP > 10, 10, siZFC3H1_rel_siGFP),
         siZFC3H1_rel_siGFP = ifelse(siZFC3H1_rel_siGFP < -10, -10, siZFC3H1_rel_siGFP)) %>%
  ggplot(., aes(x=siZCCHC8_rel_siGFP, y=siZFC3H1_rel_siGFP, color=class)) +
  geom_point(size=.1, alpha=.1) +
  facet_wrap(~libtype, scales='free') +
  geom_hline(yintercept=0, linetype=2, size=.1) +
  geom_vline(xintercept=0, linetype=2, size=.1) +
  theme_bw() +
  theme(panel.grid = element_blank())
```



#### violins

```{r}
res_df %<>%
  tidyr::separate(libtype, c('PAP', 'fraction'), sep=' ') %>%
  mutate(comparison = factor(comparison, levels=c('siRRP40_rel_siGFP', 'siZCCHC8_rel_siGFP', 'siZFC3H1_rel_siGFP', 'siZCCHC8.ZFC3H1_rel_siGFP')))
```


```{r violin plot log2FC class comparisons, width=3, height=2.5}
color_set <- c('#D42027', '#186734', '#0099CC', '#2C3A96')

res_df %>%
  filter(class != 'NOT_SIG') %>%
  ggplot(., aes(x=comparison, y=log2FoldChange, fill=comparison)) +
  geom_violin(scale = 'width') +
  geom_boxplot(fill='gray', width=.2, outlier.shape=NA) +
  coord_cartesian(ylim=c(-4,8)) +
  geom_hline(yintercept = 0, color='orange', linetype=2, size=.25) +
  facet_grid(fraction+PAP~class) +
  scale_fill_manual(values=color_set) +
  theme_bw() +
  theme(axis.text.x=element_blank(),
        axis.title.x=element_blank(),
        axis.text.y=element_text(size=6, color='black'),
        strip.text = element_text(size=6, color='black'),
        axis.title.y=element_text(size=6, color='black'),
        panel.grid=element_blank(),
        legend.title = element_text(size=6),
        legend.key.size = unit(6,'points'),
        legend.text = element_text(size=6))
```

```{r violin plot log2FC class comparisons without dblKD, width=3, height=2.5}
res_df %>%
  filter(class != 'NOT_SIG', comparison != 'siZCCHC8.ZFC3H1_rel_siGFP') %>%
  ggplot(., aes(x=comparison, y=log2FoldChange, fill=comparison)) +
  geom_violin(scale = 'width') +
  geom_boxplot(fill='gray', width=.2, outlier.shape=NA) +
  coord_cartesian(ylim=c(-4,8)) +
  geom_hline(yintercept = 0, color='orange', linetype=2, size=.25) +
  facet_grid(fraction+PAP~class) +
  scale_fill_manual(values=color_set) +
  theme_bw() +
  theme(axis.text.x=element_blank(),
        axis.title.x=element_blank(),
        axis.text.y=element_text(size=6, color='black'),
        strip.text = element_text(size=6, color='black'),
        axis.title.y=element_text(size=6, color='black'),
        panel.grid=element_blank(),
        legend.title = element_text(size=6),
        legend.key.size = unit(6,'points'),
        legend.text = element_text(size=6))
```


#### mean medians and wilcox p value of log2FC between kds


dbl vs Z8
```{r}
res_df %>% 
  group_by(class, PAP, fraction, comparison) %>%
  summarize(median_LFC=median(log2FoldChange),
            mean_LFC=mean(log2FoldChange)) %>%
  kable
```


dbl vs Z8
```{r}
res_df %>% group_by(class, PAP, fraction) %>%
  do(tidy(wilcox.test(.$log2FoldChange[grepl('^siZCCHC8.ZFC3H1', .$comparison)], .$log2FoldChange[grepl('^siZCCHC8_rel', .$comparison)])))
```

dbl vs Z1
```{r}
res_df %>% group_by(class, PAP, fraction) %>%
  do(tidy(wilcox.test(.$log2FoldChange[grepl('^siZCCHC8.ZFC3H1', .$comparison)], .$log2FoldChange[grepl('^siZFC3H1_rel', .$comparison)])))
```


#### heatmaps

```{r heatmap plot log2FC class comparisons}
res_df %>%
  filter(class != 'NOT_SIG', comparison != 'siZCCHC8.ZFC3H1_rel_siGFP') %>%
  mutate(log2FoldChange = ifelse(log2FoldChange > 10, 10, log2FoldChange),
         log2FoldChange = ifelse(log2FoldChange < -10, -10, log2FoldChange)) %>%
  ggplot(., aes(x=comparison, y=row, fill=log2FoldChange)) +
  geom_tile() +
  facet_grid(class~fraction+PAP, scales='free', space='free') +
  scale_fill_gradient2(low='firebrick4', mid='white', high = 'navyblue', midpoint=0) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45, hjust=1),
        axis.text.y=element_blank(),
        panel.grid=element_blank())
```



## prepare and save genomic location

bed files
```{r, eval=FALSE}
bed_plus="/Volumes/GenomeDK/faststorage/Guifen/Lexogen_July_Sept_Nov/Singles_clusters/Singles_union_filtgm5_merged25_filtg20_with_center_plus.bed"
bed_minus="/Volumes/GenomeDK/faststorage/Guifen/Lexogen_July_Sept_Nov/Singles_clusters/Singles_union_filtgm5_merged25_filtg20_with_center_minus.bed"

plus_anno <- rtracklayer::import(bed_plus)
minus_anno <- rtracklayer::import(bed_minus)

(clusters <- c(plus_anno, minus_anno))
```


add to classification file

```{r, eval=FALSE}
cluster_class_map <- cluster_classes$class
names(cluster_class_map) <- cluster_classes$cluster
```

```{r, eval=FALSE}
clusters$class <- cluster_class_map[clusters$name]
clusters$name <- paste0(clusters$name, ':', clusters$class)
```

save this
```{r, eval=FALSE}
rtracklayer::export.bed(clusters,con = '../../data/clusters.bed')
```


## load clusters and annotation
```{r, eval=TRUE}
(clusters <- rtracklayer::import('../../data/clusters.bed'))
```

```{r}
clusters$class <- sub('.*:', '', clusters$name)
clusters$cluster <- sub(':.*', '', clusters$name)
clusters$center <- sapply(clusters$name, function(n) as.integer(str_split(n, ':', simplify=T)[2]))
clusters
```

```{r}
(clusters_tbl <- data.frame(clusters) %>% 
  tbl_df)
```


#### clusters per class counts

n:
```{r}
table(clusters_tbl$class)
```

fraction:
```{r}
table(clusters_tbl$class)/sum(table(clusters_tbl$class))
```

#### amount of genomic positions in the various categories

total bp in clusters:
```{r}
(total_cluster_bps <- sum(clusters_tbl$width))
```

total bp per class

```{r}
clusters_tbl %>%
  group_by(class) %>%
  summarize(total_bp = sum(width)) %>%
  mutate(freq_in_class = total_bp/total_cluster_bps)
```


```{r stacked barplot bps per class}
clusters_tbl %>%
  group_by(class) %>%
  summarize(total_bp = sum(width)) %>%
  mutate(freq_in_class = total_bp/total_cluster_bps,
         class = factor(class, levels=c('PAXT', 'BOTH', 'NEXT', 'NOT_SIG'))) %>%
  ggplot(., aes(x=1, y=total_bp, fill=class)) +
  geom_bar(stat='identity')
```

#### amount of total signal in the various categories

need to get signal per cluster.
obviously depends a lot on KD and library type, simply do this for all...
```{r}
load('../../data/count_mat_RRP40Z1Z8clusters.RData', verbose=TRUE)
glimpse(count_mat)
```

normalize raw reads
```{r}
(nfs <- read_tsv('../../data/last_exon_sizeFactors.table') %>%
  mutate(full_lib_name = paste0(series, '_', lib_name)) %>%
  dplyr::select(full_lib_name, geom_mean_sf))
```


```{r}
(norm_cnt_df <- count_mat %>%
  data.frame %>% 
  rownames_to_column('cluster') %>%
  gather(full_lib_name, cnt, -cluster) %>%
  tbl_df %>%
  left_join(., nfs) %>%
  mutate(norm_cnt = cnt/geom_mean_sf) %>%
  dplyr::select(cluster, full_lib_name, norm_cnt) %>%
  tidyr::separate(full_lib_name, c('series', 'siRNA', 'PAP', 'fraction', 'rep'), sep='_') %>%
  group_by(cluster, siRNA, fraction, PAP) %>%
  summarise(norm_cnt = mean(norm_cnt)))
```

```{r}
(norm_cnt_df %<>% left_join(., cluster_classes))
```

```{r}
save(norm_cnt_df, file = '../../data/norm_cluster_counts.RData')
```

```{r}
distinct(norm_cnt_df, cluster, class) %$%
  class %>% 
  table
```


```{r}
(total_cluster_signal <- norm_cnt_df %>%
   group_by(siRNA, fraction, PAP) %>%
   summarize(total_incluster_norm_cnt = sum(norm_cnt)))
```


total bp per class

```{r}
(signal_per_class <- norm_cnt_df %>%
   group_by(class, siRNA, fraction, PAP) %>%
   summarize(total_norm_cnt_per_class = sum(norm_cnt)) %>%
  left_join(., total_cluster_signal) %>%
   mutate(freq_in_class = total_norm_cnt_per_class/total_incluster_norm_cnt))
```


```{r}
signal_per_class %>%
  filter(siRNA == 'siGFP')
```

```{r}
signal_per_class %>%
  filter(siRNA == 'siRRP40')
```

```{r}
signal_per_class %>%
  filter(siRNA == 'siZCCHC8')
```

```{r}
signal_per_class %>%
  filter(siRNA == 'siZFC3H1')
```

```{r}
signal_per_class %>%
  filter(siRNA == 'siZCCHC8.ZFC3H1')
```


```{r stacked barplot total signal per class}
signal_per_class %>%
  ungroup %>%
  mutate(class = factor(class, levels=c('PAXT', 'BOTH', 'NEXT', 'NOT_SIG'))) %>%
  ggplot(., aes(x=siRNA, y=total_norm_cnt_per_class, fill=class)) +
  geom_bar(stat='identity') +
  facet_grid(fraction~PAP) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle=45, hjust=1))
```


#### add signal outside clusters

```{r}
(total_reads_per_lib <- read_tsv('../../data/Raw_reads_per_lib.txt'))
```



```{r}
(norm_total_reads_per_lib <- total_reads_per_lib %>%
  mutate(fraction = sub('put', '', fraction),
         siRNA = sub('\\+', '\\.', siRNA),
         full_lib_name = paste0(series, '_', siRNA, '_', PAP, '_', fraction), 
         full_lib_name = ifelse(is.na(replicate), full_lib_name, paste0(full_lib_name, '_', replicate))) %>%
  dplyr::select(full_lib_name, reads) %>%
  left_join(., nfs) %>%
  mutate(norm_genome_reads = reads/geom_mean_sf) %>%
  tidyr::separate(full_lib_name, c('series', 'siRNA', 'PAP', 'fraction', 'replicate'), sep='_') %>%
  group_by(siRNA, PAP, fraction) %>%
  summarize(total_genome_norm_cnt = mean(norm_genome_reads)))
```


```{r norm reads in genome vs in cluster}
left_join(total_cluster_signal, norm_total_reads_per_lib) %>%
  gather(in_cluster, norm_cnt, -siRNA, -fraction, -PAP) %>%
  ggplot(., aes(x=siRNA, y=norm_cnt, fill=in_cluster)) +
  geom_bar(stat='identity', position='dodge') +
  facet_grid(.~PAP+fraction) +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=1),
        panel.grid=element_blank())
```


```{r}
(not_cluster <- left_join(total_cluster_signal, norm_total_reads_per_lib) %>%
  mutate(total_norm_cnt_per_class = total_genome_norm_cnt-total_incluster_norm_cnt, 
         class = 'outside_cluster') %>%
  dplyr::select(class, siRNA, PAP, fraction, total_genome_norm_cnt, total_norm_cnt_per_class))
```


```{r}
signal_per_class_vs_genome <- left_join(signal_per_class, norm_total_reads_per_lib) %>%
  dplyr::select(class, siRNA, PAP, fraction, total_genome_norm_cnt, total_norm_cnt_per_class) %>%
  bind_rows(., not_cluster) %>%
  mutate(freq_in_class = total_norm_cnt_per_class/total_genome_norm_cnt)
```


```{r stacked barplot total signal per class vs genome}
signal_per_class_vs_genome %>%
  ungroup %>%
  mutate(class = factor(class, levels=c('PAXT', 'BOTH', 'NEXT', 'NOT_SIG', 'outside_cluster'))) %>%
  ggplot(., aes(x=siRNA, y=total_norm_cnt_per_class, fill=class)) +
  geom_bar(stat='identity') +
  facet_grid(fraction~PAP) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle=45, hjust=1))
```

```{r stacked barplot freq per class vs genome}
signal_per_class_vs_genome %>%
  ungroup %>%
  mutate(class = factor(class, levels=c('PAXT', 'BOTH', 'NEXT', 'NOT_SIG', 'outside_cluster'))) %>%
  ggplot(., aes(x=siRNA, y=freq_in_class, fill=class)) +
  geom_bar(stat='identity') +
  facet_grid(fraction~PAP) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle=45, hjust=1))
```

```{r}
signal_per_class_vs_genome %>%
  dplyr::rename(genome = total_genome_norm_cnt,
                in_class = total_norm_cnt_per_class) %>%
  filter(siRNA == 'siGFP')
```

```{r}
signal_per_class_vs_genome %>%
  dplyr::rename(genome = total_genome_norm_cnt,
                in_class = total_norm_cnt_per_class) %>%
  filter(siRNA == 'siRRP40')
```

```{r}
signal_per_class_vs_genome %>%
  dplyr::rename(genome = total_genome_norm_cnt,
                in_class = total_norm_cnt_per_class) %>%
  filter(siRNA == 'siZCCHC8')
```

```{r}
signal_per_class %>%
  dplyr::rename(genome = total_genome_norm_cnt,
                in_class = total_norm_cnt_per_class) %>%
  filter(siRNA == 'siZFC3H1')
```

```{r}
signal_per_class_vs_genome %>%
  dplyr::rename(genome = total_genome_norm_cnt,
                in_class = total_norm_cnt_per_class) %>%
  filter(siRNA == 'siZCCHC8.ZFC3H1')
```


```{r}
sessionInfo()
```
