---
title: 'CLIPdb on cluster classes'
author: "Manfred Schmid"
output: 
  pdf_document:
    toc: true 
    toc_depth: 3
    fig_caption: true
editor_options: 
  chunk_output_type: console
---

`r format(Sys.time(), "%d %B, %Y; %R")`

## Setup


```{r setup, echo=TRUE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, 
                      fig.path='../../Figures/cluster_CLIPdb/', 
                      dev='pdf', 
                      echo=TRUE, warning=FALSE, message=FALSE, 
                      error=TRUE)
```


```{r load packages, echo=T, warning=F, message=F}
suppressWarnings(library('tidyverse'))
suppressWarnings(library('magrittr'))
suppressWarnings(library('knitr'))
suppressWarnings(library('broom'))
suppressWarnings(library('rtracklayer'))
```



## clusters new

```{r}
(clusters <- rtracklayer::import('../../data/clusters.bed'))
```

```{r}
clusters$class <- sub('.*:', '', clusters$name)
clusters$name <- sub(':BOTH|:NOT_SIG|:NEXT|:PAXT', '', clusters$name)
clusters$summit <- as.integer(sub('.*:', '', clusters$name))+1
#clusters$summit <- ifelse(strand(clusters)=="+", clusters$summit+1, clusters$summit)

clusters
```


```{r}
clusters_summitpm100 <- clusters[clusters$class != 'BOTH' & clusters$class != 'NOT_SIG',]

start(clusters_summitpm100) <- clusters_summitpm100$summit - 100
end(clusters_summitpm100) <- clusters_summitpm100$summit + 100


clusters_summitpm100_plus <- clusters_summitpm100[strand(clusters_summitpm100)=='+',]

clusters_summitpm100_minus <- clusters_summitpm100[strand(clusters_summitpm100)=='-',]
```



```{r}
CLIPdb_bw_path <- '/Volumes/GenomeDK/THJ_common/faststorage/data/Human/CLIPdb/hg38_bw/'

bw_fnames_base <- dir(CLIPdb_bw_path) %>% keep(grepl('plus.bw', .)) %>% sub('_plus.bw', '', .)
```


```{r}
get_bw_sum <- function(bw, ranges){
  
  bw_file <- rtracklayer::BigWigFile(bw)

  seqlevels(ranges, pruning.mode="coarse") <- seqnames(seqinfo(bw_file))
  
    values <- lapply(rtracklayer::import(bw_file, which=ranges, as = 'NumericList'),
    sum) %>% unlist
  
  data.frame(cluster=ranges$name,
             class=ranges$class,
             value=values)
}
```

```{r}
get_data_for_CLIPdb <- function(bw_base){
  print(paste0(' doing: ', bw_base))
  plus_bw <- paste0(CLIPdb_bw_path,bw_base,'_plus.bw')
  minus_bw <- paste0(CLIPdb_bw_path,bw_base,'_minus.bw')
  ex_plus <- get_bw_sum(plus_bw, 
                      clusters_summitpm100_plus)
           

  ex_minus <- get_bw_sum(minus_bw, 
                      clusters_summitpm100_minus)

  bind_rows(ex_plus,ex_minus) %>% tbl_df %>% mutate(experiment = bw_base)
}

```


#### example WDR33


```{r}
(wdr33_base <- bw_fnames_base %>% keep(grepl('WDR33', .)))
```

```{r}
wdr33_meta <- get_data_for_CLIPdb(wdr33_base)
```


```{r}
ggplot(wdr33_meta, aes(x=class, y=log10(value+1), fill=class)) +
  geom_violin()
```


```{r}
wdr33_meta %>%
  do(tidy(t.test(.$value~.$class)))
```



#### get all metavalues


some bw files here are corrupt and useless, this can be inferred from their size:

```{r}
sapply(bw_fnames_base, function(base) file.size(paste0(CLIPdb_bw_path, base, '_plus.bw'))) %>%
  sort
```

--> DGCR8_eCLIP and HNRNPU_eCLIP are extremely small and will be removed

```{r}
bw_fnames_base %<>%
  keep(!grepl('DGCR8_eCLIP', .) & !grepl('HNRNPU_eCLIP', .))
```


```{r, eval=FALSE}
(meta <- lapply(bw_fnames_base, get_data_for_CLIPdb) %>% bind_rows)
```


```{r, eval=FALSE}
save(meta, file = '../../data/CLIPdb_summitpm200_sums.RData')
```


```{r}
load('../../data/CLIPdb_summitpm200_sums.RData', verbose=T)
meta
```

```{r}
(meta_ttests <- meta %>%
  group_by(experiment) %>%
  do(tidy(t.test(.$value~.$class))))
```


```{r}
meta_ttests %>% ungroup %>% arrange(p.value)
```


## subsets by annotation


```{r}
load('../../data/clusters_with_cat_and_geneanno.RData', verbose=T)
overlaps_df
```


```{r}
table(overlaps_df$gene_biotype)
```


#### not overlapping splice junctions

splice junctions are relatively abundant in NEXT subset. This is a specific substrate class not relevant for the present manuscript. Since these bias for enrichment for splice factors, which are not relevant here, we focus on non-splice junction clusters.

```{r}
table(overlaps_df$junction_type)
```


```{r}
nonSJ_ids <- overlaps_df %>%
  filter(junction_type == 'none') %$% 
  name %>%
  sub(':.*', '', .)
```


```{r}
nonSJ_meta <- filter(meta, sub(':.*', '', cluster) %in% nonSJ_ids)
```



```{r}
(nonSJ_meta_wilcoxtests <- nonSJ_meta %>%
  group_by(experiment) %>%
  do(tidy(wilcox.test(.$value~.$class))))
```

```{r}
(nonSJ_meta_ttests <- nonSJ_meta %>%
  group_by(experiment) %>%
  do(tidy(t.test(log2(.$value+1)~.$class))))
```

```{r}
levels(nonSJ_meta$class)
```

```{r}
nonSJ_meta_wilcoxtests %>% ungroup %>% arrange(p.value)
```

```{r}
nonSJ_meta_ttests %>% ungroup %>% arrange(-estimate)
```

```{r}
nonSJ_meta_ttests %>% ungroup %>% arrange(estimate)

```

```{r}
nonSJ_meta_ttests %>% ungroup %>% arrange(p.value)
```

```{r}
nonSJ_tt_wilcox_p <- nonSJ_meta_wilcoxtests %>%
  dplyr::select(experiment, p.value) %>%
  dplyr::rename(wilcox_p = p.value) %>%
  ungroup %>%
  mutate(wilcox_padj = p.adjust(wilcox_p)) %>%
  left_join(nonSJ_meta_ttests, .)
```

```{r non SJ PAXT vs NEXT log2FC with wilcox p value coloring}
ggplot(nonSJ_tt_wilcox_p,
       aes(x=reorder(experiment, -estimate), y=-estimate, fill=-log10(wilcox_padj))) +
         geom_bar(stat='identity', position='dodge') +
  scale_fill_gradient(low='lightgray', high='firebrick4') +
  theme(axis.text.x = element_text(angle=45, hjust=1, size=1),
        panel.grid=element_blank(),
        panel.background = element_blank())
```


```{r non SJ PAXT vs NEXT log2FC with wilcox p value coloring top hits}
nonSJ_tt_wilcox_p %>%
  filter(estimate < -.1) %>%
  ggplot(.,
       aes(x=reorder(experiment, -estimate), y=-estimate, fill=-log10(wilcox_padj))) +
         geom_bar(stat='identity', position='dodge') +
  scale_fill_gradient(low='lightgray', high='firebrick4') +
  theme(axis.text.x = element_text(angle=45, hjust=1, size=6),
        panel.background=element_blank(),
        panel.grid=element_blank())
```





#### for only intergenic or unstable types

```{r}
intergenics <- overlaps_df %>%
  filter(gene_biotype == 'intergenic' |
           grepl('antisense', gene_biotype) |
           grepl('lncRNA', gene_biotype) |
           grepl('pseudogene', gene_biotype))
```

```{r}
table(intergenics$gene_biotype)
```

```{r}
intergenic_cluster_ids <- sub(':.*', '', intergenics$name)
```


```{r}
intergenic_meta <- filter(meta, sub(':.*', '', cluster) %in% intergenic_cluster_ids)
```



```{r}
(intergenic_meta_wilcoxtests <- intergenic_meta %>%
  group_by(experiment) %>%
  do(tidy(wilcox.test(.$value~.$class))))
```

```{r}
(intergenic_meta_ttests <- intergenic_meta %>%
  group_by(experiment) %>%
  do(tidy(t.test(log2(.$value+1)~.$class))))
```

```{r}
levels(intergenic_meta$class)
```

```{r}
intergenic_meta_wilcoxtests %>% ungroup %>% arrange(p.value)
```

```{r}
intergenic_meta_ttests %>% ungroup %>% arrange(-estimate)
```

```{r}
intergenic_meta_ttests %>% ungroup %>% arrange(estimate)

```

```{r}
intergenic_meta_ttests %>% ungroup %>% arrange(p.value)
```

```{r}
intergenic_tt_wilcox_p <- intergenic_meta_wilcoxtests %>%
  dplyr::select(experiment, p.value) %>%
  dplyr::rename(wilcox_p = p.value) %>%
  ungroup %>%
  mutate(wilcox_padj = p.adjust(wilcox_p)) %>%
  left_join(intergenic_meta_ttests, .)
```

```{r intergenic PAXT vs NEXT log2FC with wilcox p value coloring}
ggplot(intergenic_tt_wilcox_p,
       aes(x=reorder(experiment, -estimate), y=-estimate, fill=-log10(wilcox_padj))) +
         geom_bar(stat='identity', position='dodge') +
  scale_fill_gradient(low='lightgray', high='firebrick4') +
  theme(axis.text.x = element_text(angle=45, hjust=1, size=1))
```


```{r intergenic PAXT vs NEXT log2FC with wilcox p value coloring top hits}
intergenic_tt_wilcox_p %>%
  filter(estimate < -.15) %>%
  ggplot(.,
       aes(x=reorder(experiment, -estimate), y=-estimate, fill=-log10(wilcox_padj))) +
         geom_bar(stat='identity', position='dodge') +
  scale_fill_gradient(low='lightgray', high='firebrick4') +
  theme(axis.text.x = element_text(angle=45, hjust=1, size=6),
        panel.background=element_blank(),
        panel.grid=element_blank())
```



# metagene profiles

```{r}
get_bw_values <- function(bw, ranges){
  
  bw_file <- rtracklayer::BigWigFile(bw)

  seqlevels(ranges, pruning.mode="coarse") <- seqnames(seqinfo(bw_file))
  
    values <- rtracklayer::import(bw_file, which=ranges, as = 'NumericList')
  
  df <- data.frame(as.matrix(values))
  df$cluster <- ranges$name
  df$class <- ranges$class
  
  df %<>%
    gather(pos, value, -cluster, -class) %>%
    mutate(pos = as.integer(sub('X','',pos)))
  
  tbl_df(df)
}
```


```{r}
get_data_for_CLIPdb_profile <- function(bw_base, plus_ranges, minus_ranges){
  print(paste0(' doing: ', bw_base))
  plus_bw <- paste0(CLIPdb_bw_path,bw_base,'_plus.bw')
  minus_bw <- paste0(CLIPdb_bw_path,bw_base,'_minus.bw')
  ex_plus <- get_bw_values(plus_bw, 
                      plus_ranges)
           

  ex_minus <- get_bw_values(minus_bw, 
                      minus_ranges)

  bind_rows(ex_plus,ex_minus) %>%  mutate(experiment = bw_base)
}

```


#### for top nonSJ hits

```{r}
(exp_sig <- nonSJ_tt_wilcox_p %>%
  filter(estimate < -.1) %$% 
  experiment)
```


```{r}
clusters_summitpm1kb <- clusters[clusters$class != 'BOTH' & clusters$class != 'NOT_SIG' & sub(':.*', '', clusters$name) %in% nonSJ_ids,]

start(clusters_summitpm1kb) <- clusters_summitpm1kb$summit - 1000
end(clusters_summitpm1kb) <- clusters_summitpm1kb$summit + 1000


clusters_summitpm1kb_plus <- clusters_summitpm1kb[strand(clusters_summitpm1kb)=='+',]

clusters_summitpm1kb_minus <- clusters_summitpm1kb[strand(clusters_summitpm1kb)=='-',]
```

```{r}
(meta_profiles <- lapply(exp_sig[1], function(exp) get_data_for_CLIPdb_profile(exp, clusters_summitpm1kb_plus, clusters_summitpm1kb_minus) %>%
  mutate(pos = ifelse(grepl('^cp',cluster), pos-1001, 1001-pos)) %>%
  group_by(pos, class, experiment) %>%
  summarize(total_log_value=mean(log2(value+.001)))) %>% 
   bind_rows)
```


```{r}
meta_profile_avg <- meta_profiles %>%
  mutate(pos = ifelse(grepl('^cp',cluster), pos-1001, 1001-pos)) %>%
  group_by(pos, class, experiment) %>%
  summarize(total_log_value=mean(log2(value+.001)))
```

```{r non SJ metagene profiles}
ggplot(meta_profile_avg,
       aes(x=pos, y=total_log_value, color=class)) +
  geom_line() +
  facet_wrap(~experiment)
```


#### for top hits intergenics

```{r}
(exp_sig <- intergenic_tt_wilcox_p %>%
  filter(estimate < -.15) %$% 
  experiment)
```


```{r}
clusters_summitpm1kb <- clusters[clusters$class != 'BOTH' & clusters$class != 'NOT_SIG' & sub(':.*', '', clusters$name) %in% intergenic_cluster_ids,]

start(clusters_summitpm1kb) <- clusters_summitpm1kb$summit - 1000
end(clusters_summitpm1kb) <- clusters_summitpm1kb$summit + 1000


clusters_summitpm1kb_plus <- clusters_summitpm1kb[strand(clusters_summitpm1kb)=='+',]

clusters_summitpm1kb_minus <- clusters_summitpm1kb[strand(clusters_summitpm1kb)=='-',]
```

```{r}
(meta_profiles <- lapply(exp_sig, function(exp) get_data_for_CLIPdb_profile(exp, clusters_summitpm1kb_plus, clusters_summitpm1kb_minus)) %>% bind_rows)
```


```{r}
meta_profile_avg <- meta_profiles %>%
  mutate(pos = ifelse(grepl('^cp',cluster), pos-1001, 1001-pos)) %>%
  group_by(pos, class, experiment) %>%
  summarize(total_log_value=mean(log2(value+.001)))
```

```{r}
ggplot(meta_profile_avg,
       aes(x=pos, y=total_log_value, color=class)) +
  geom_line() +
  facet_wrap(~experiment)
```


```{r}
intergenics <- overlaps_df %>%
  filter(gene_biotype == 'intergenic')
```

```{r}
table(intergenics$gene_biotype)
```

```{r}
intergenic_cluster_ids <- intergenics$name
```

```{r}
inter_meta <- filter(meta_profiles, cluster %in% intergenic_cluster_ids)
```


```{r}
inter_meta_profile_avg <- inter_meta %>%
  mutate(pos = ifelse(grepl('^cp',cluster), pos-1001, 1001-pos)) %>%
  group_by(pos, class, experiment) %>%
  summarize(total_log_value=mean(log2(value+.001)))
```

```{r intergenic metagene profiles}
ggplot(inter_meta_profile_avg,
       aes(x=pos, y=total_log_value, color=class)) +
  geom_line() +
  facet_wrap(~experiment)
```


# sessionInfo
```{r}
sessionInfo()
```
